---
title: "Econometrics II Assignment"
---

Working Directory
```{r}
setwd("C:/Users/Marco Hassan/Desktop/Econometrics II/Assignment/Data")
```

Libararies
```{r}
library(boot)
library(fields)
library(robustbase)
library(ggplot2)
library(gmm)
library(moments)
library(AER)
```

===
Simple OLS replication
===

Dataset
```{r}
library(haven)
X2_OLS <- read_dta("Data/2. OLS.dta")
```

Select Base sample data
```{r}
base <-  X2_OLS[!is.na(X2_OLS$baseco >= 1), 1:8]
```

Regression
```{r}
lm1 <- lm(base$logpgp95 ~  base$avexpr + base$lat_abst + base$africa + base$africa)

lm2 <- lm(X2_OLS$logpgp95 ~ X2_OLS$avexpr)

plot(lm1)
```


===
Bootstrapping intuition
===

Create data frame
```{r}
t1 = as.data.frame(cbind(base$logpgp95, base$avexpr, base$lat_abst, base$africa, base$asia))
names(t1) <- colnames(base[c(5, 4, 3, 2, 7)])

t2 = as.data.frame(cbind(X2_OLS$logpgp95, X2_OLS$avexpr))
names(t2) <- colnames(base[c(5, 4)])
```

Sample from the given data frame
```{r}
nb = 5000
bet = NULL ; n = NROW(t1)
for (i in 1:nb){
  unifnum = sample(c(1:n),n,replace = T)	
  bet[i] = lm(t1[unifnum,1]~t1[unifnum,2] + t1[unifnum,3]+ t1[unifnum,4]+ t1[unifnum,5])$coef[2]
}

bet2=NULL; n2 = NROW(t2)
for (i in 1:nb){
  unifnum = sample(c(1:n2),n2,replace = T)	
  bet2[i] = lm(t2[unifnum,1]~t2[unifnum,2])$coef[2]
}
```

--> Not working: not converging

bet2.2 = NULL; n2.2 = NROW(t1)
for (i in 1:nb){
  unifnum = sample(c(1:n2.2),n2.2,replace = T)	
  LL <- function(beta0, beta1, mu, sigma) {
    R = t1[unifnum, 1] - t2[unifnum,2] * beta1 - beta0
    R = suppressWarnings(dnorm(R, mu, sigma))
    -sum(log(R))}
    
    bet2.2[i] <- coef(mle(LL, start = list(beta0 = 5, beta1 = 0.5, mu = 0.08,
                                           sigma=.7)))[2]
}


Plot results
```{r}
histo <- function(bet, lm1){
  lwd1 = 2.5
  
  den = density(bet)
  bethat <- bquote(Histogram~of ~bold(widehat(beta))) 
  h = hist(bet, breaks = 100, freq = NULL, probability = F, xlab =
             expression(bold(widehat(beta))), cex.lab = 1.6, main =   bethat )
  xline(mean(bet), lty = 6, col = "blue", lwd = lwd1) ; xline(lm1$coef[2], lty = 6,
                                                              col = 2, lwd = lwd1)
  xline(mean(bet)+1.96*summary(lm1)$coef[2,2],lty = 6, col = 2, lwd = lwd1)
  xline(mean(bet)-1.96*summary(lm1)$coef[2,2], lty= 6, col = 2, lwd = lwd1)
  xline(quantile(bet,.025), col = "blue", lty = 6, lwd = lwd1) ;
  xline(quantile(bet,.975), col = "blue", lty=6, lwd = lwd1)
  
  xfit<-seq(min(bet),max(bet),length=length(bet))
  yfit<-dnorm(xfit,mean=mean(bet),sd=sd(bet))
  yfit <- yfit*diff(h$mids[1:2])*length(bet)
  lines(xfit, yfit, col="blue", lwd=2)
  yfit2<-dnorm(xfit,mean=lm1$coef[2],sd=summary(lm1)$coef[2,2])
  yfit2 <- yfit2*diff(h$mids[1:2])*length(bet)
  lines(xfit, yfit2, col=2, lwd=2)
  legend("topleft", legend=c("empirical denisty statistics", "normal denisty
                             statistics"),
       col=c("blue","red"), lty= 6, cex=0.8)
}
```

Histogram base sample
```{r}
histo(bet, lm1)
histo(bet2, lm2)
```

Try model based resampling
```{r}

t3 <- as.data.frame(na.omit(cbind(X2_OLS$logpgp95, X2_OLS$avexpr)))
colnames(t3) <-  cbind("Log", "Aver")

anBase <- anova(lm(t3$Log ~ t3$Aver))
Fbase <- anBase["t3$Aver", "F value"]
pBase <- anBase["t3$AVer", "Pr(>F)"]

#H0-Modell in Basisstichprobe anpassen
fit0 <- lm(t3$Log  ~ 1) # Modellanpassung
E <- residuals(fit0) # ursprüngliche Residuen
Er <- E / sqrt(1-hatvalues(fit0)) # reskalierte Residuen
Yhat <- fitted(fit0) # ursprüngliche Vorhersage
# ANOVA-Parameter für ursprüngliche Gruppenzugehörigkeiten und Y*
getAnova <- function(dat, idx) {
 Ystar <- Yhat + Er[idx] # Resample-AV Y* = Y^ + E*
 anBS <- anova(lm(Ystar ~ t3$Aver, data=dat)) # ANOVA Resample-AV
 anBS["t3$Aver", "F value"] # F*-Wert Replikation
}

nR <- 5000 # Anzahl Replikationen
bsAnova <- boot(t3, statistic=getAnova, R=nR)

## F* aus Bootstrap-Replikationen
Fstar <- bsAnova$t
# numerische Toleranz wg. Gleitkomma-Arithmetik
tol <- .Machine$double.eps^0.5
# Fstar größer Basis-F ODER Fstar ungefähr gleich Basis-F?
FsIsGEQ <- (Fstar > Fbase) | (abs(Fstar - Fbase) < tol)
# p-Wert: Anteil der mindestens so extremen F* wie Basis F
(pValBS <- (sum(FsIsGEQ) + 1) / (length(Fstar) + 1))
# Grafik: kumulierte relative Häufigkeiten von F*

plot(Fstar, ecdf(Fstar)(Fstar), col="gray60", pch=1,
 xlab="f* bzw. f", ylab="P(F <= f)",
 main="F*: Kumulierte rel. Häufigkeiten und Verteilungsfunktion")
# theoretische Verteilungsfunktion von t und Legende
curve(pf(x, 1, 110), lwd=2, add=TRUE)
 legend(x="topleft", lty=c(NA, 1), pch=c(1, NA), lwd=c(2, 2),
 col=c("gray60", "black"), legend=c("F*", "F"))


```






Constructed test on real normal distributed variable 
```{r}
eps <- rnorm(NROW(t1), 0, sd(base$avexpr))
fakey <- predict(lm1) +eps
lm_fake = lm(fakey~base$avexpr)
fakebet = NULL
for (i in 1:nb){
  unifnum = sample(c(1:n),n,replace = T)	
  fakebet[i] = lm(fakey[unifnum]~base$avexpr[unifnum])$coef[2]
}

histo(fakebet, lm_fake)
```

QQ-plot
```{r}
## QQ-plots
qqnorm(bet); qqline(bet, col ="red", lty =2, lwd = 2)
```

Confidence limits
```{r}
confidence <- function(bet, lm1){
  bias <- mean(bet) - lm1$coefficients[2]
  
  boot_coorected_estim <- lm1$coefficients[2] - bias
  
  ## Assuming normality, the approximate 95% confidence interval is given by CVd ± 1.96pVar(bootstrap)
  
  Low <- lm1$coefficients[2] - bias - 1.96*sqrt(var(bet))
  High <- lm1$coefficients[2] - bias + 1.96*sqrt(var(bet))
  return(c("bias"=bias, "corrected estimate"=boot_coorected_estim, "low CI"= Low,
           "High CI" = High))
}
```

Confidence results
```{r}
confidence(bet, lm1)
summary(lm1)
confidence(bet2, lm2)
```

Using boot package
```{r}
beta_function <- function(dat, i, func){
  resampled <- dat[i, ]
  return(lm(resampled[,1]~resampled[,2]+ resampled[,3]+ resampled[,4]+ resampled[,5], data = resampled)$coef[2]) 
}
```

Bootstrapped results
```{r}
bootbet <- boot(dat=t1, statistic=beta_function, R=5000) 
bootbet
```


Confidence intervals
```{r}
boot.ci(boot.out = bootbet, type = c("norm", "basic", "perc", "bca"))

jack.after.boot(bootbet, main = "Jackknife after Boot in  simple OLS")
```

===
Plots of IV approach
===

Import Dataset and select relevant base sample
```{r}
library(haven)
X4_IV_regression_of_Log_GDP_per_capita <- read_dta("Data/4. IV regression of Log_GDP per capita.dta")

base2 <- X4_IV_regression_of_Log_GDP_per_capita[!is.na(X4_IV_regression_of_Log_GDP_per_capita$baseco >= 1), 1:9]
```



===
Plotting results
===
Plot reduced form
```{r}
c<- ggplot(base2, aes(x=base2$logem4, y=base2$logpgp95)) + 
  geom_point(color='black', size = 2, alpha =0.8) + 
  geom_smooth(method=lm, color='blue')+
  xlab("Log Settler Mortality")+ ylab("Log GDP")+ ggtitle("Reduced form regression")+
  theme_bw()
```


Plot First Stage
```{r}
a <- ggplot(base2, aes(x=base2$logem4, y=base2$avexpr)) + 
  geom_point(color='black', size = 2, alpha =0.8) + 
  geom_smooth(method=lm, color='blue')+
  xlab("Log Settler Mortality")+ ylab("Average expropration rate")+ ggtitle("First stage regression")+
  theme_bw()
```

Plot second stage
```{r}
b <- ggplot(base2, aes(x=base2$avexpr, y=base2$logpgp95)) + 
  geom_point(color='black', size = 2, alpha =0.8) + 
  geom_smooth(method=lm, color='blue')+
  xlab("Average expropration rate")+ ylab("Log GDP")+ ggtitle("Second stage regression")+
  theme_bw()
```

Multiplot function
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }

  if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

Multiplot three regressions
```{r}
multiplot(plotlist = list(a,b,c), cols = 3)
```

===
Correlation Analysis
===
```{r}
help <- as.matrix(cbind(base2$africa, base2$lat_abst, base2$rich4, base2$avexpr,
                        base2$logpgp95,  base2$logem4, base2$asia))
colnames(help) <-colnames(base2[,c(-1,-9)])
cor(help)
```


===
OLS Reduced Form
===

OLS
```{r}
lm3 <- lmrob(base2$logpgp95 ~ base2$logem4 + base2$africa + base2$asia + 
            base$lat_abst)

summary(lm3)

plot.new()
layout(matrix(1:4,2,2)) 
plot(lm3)
```


===
Bootstrapped Variance
===

Using boot package
```{r}
beta_function2<- function(dat, i, func){
  resampled <- dat[i, ]
  return(lmrob(resampled[,1]~resampled[,2]+ resampled[,3]+ resampled[,4] , dat = resampled)$coef[2]) 
}
```

Create Dataframe 
```{r}
t3 <-  as.data.frame(cbind(base2$logpgp95, base2$logem4, base2$africa, base2$lat_abst))
```


Bootstrapped results
```{r}
bootbet2 <- suppressWarnings(boot(dat=t3, statistic=beta_function2, R=5000)) 
```

Histogram plot
```{r}
histo(bootbet2$t, lm3)
```

Interpretation of the shifted curve 
```{r}
jack.after.boot(bootbet2) ## without 20 and 51 much shorter confidence intervals
## Huge effect of singapore!!
```


Confidence intervals
```{r}
boot.ci(boot.out = bootbet2, type = c("norm", "basic", "perc", "bca"))
```

===
GMM
===

OLS First stage
```{r}
lm4 <- lm(base2$avexpr ~ base2$logem4 + base2$africa + base2$asia + 
            base$lat_abst)

summary(lm4)
layout(matrix(1:4, 2, 2))
plot(lm4)

```

GMM
```{r}
GMM <- gmm(base2$logpgp95 ~ base2$avexpr + base2$africa+
       base2$asia + base$lat_abst,  ~  base2$logem4 + base2$africa+
       base2$asia + base$lat_abst, type="twoStep", wmatrix = "optimal")

summary(GMM)

confint(GMM)
```

===
Bootstrapped Variance GMM
===
```{r}
t4 <- as.data.frame(cbind(base2$logpgp95, base2$avexpr, base2$africa, base2$asia,
                          base2$lat_abst,  base2$logem4))


beta_function3<- function(dat, i){
  resampled <<- dat[i, ]
  return(gmm(resampled[,1] ~ resampled[,2] + resampled[,3]+
       resampled[,4] + resampled[,5], ~ resampled[,6] + resampled[,3]+
       resampled[,4] + resampled[,5], type="twoStep", vcov = "iid", wmatrix = "optimal")$coef[2])
}

bootbet3 <- boot(t4, statistic=beta_function3, R=5000)
```


Plot bootstrapped gmm coefficients
```{r}
plot.new()
lwd1 = 2.5
 
hist(bootbet3$t, breaks = 12000, freq = NULL, probability = F, xlim=range(-6:6), 
     xlab = expression(bold(widehat(beta))), axes=T, cex.lab = 1.6, 
     main = bquote(Histogram~of ~bold(widehat(beta))~ GMM), col=rgb(0,0,1,1/3))
```

Summary statistics bootstrapped coefficients
```{r}
summary(bootbet3$t)
skewness(bootbet3$t)
kurtosis(bootbet3$t)
```

QQ-plot
```{r}
qqnorm(bootbet3$t, main="GMM Normal QQ-Plot"); qqline(bootbet3$t, col ="red", lty =2, lwd = 2)
```

Jack after boot
```{r}
jack.after.boot(bootbet3, main = "Jack-knife on Bootstrapped GMM")
```

Confidence intervals
```{r}
boot.ci(bootbet3)
bootbet3
```

===
TSLS 
===

```{r}
TSLS <- ivreg(base2$logpgp95 ~ base2$avexpr + base2$africa+
       base2$asia + base$lat_abst | base2$logem4 + base2$africa+
       base2$asia + base$lat_abst)

summary(TSLS)

confint(TSLS)
```

===
Bootstrapped variance TSLS
===
```{r}
beta_function4<-  function(dat, i){
  resampled2 <<- dat[i,]
  bet4[i] <- return(ivreg(resampled2[,1] ~ resampled2[,2] + resampled2[,3]+
       resampled2[,4] + resampled2[,5]| resampled2[,6] + resampled2[,3]+
       resampled2[,4] + resampled2[,5])$coefficients[2])
}


bootbet4 <- boot(data = t4, beta_function4, R=5000)
```

TSLS histogram
```{r}
hist(bootbet4$t, breaks = 7000, freq = NULL, probability = F, xlim=range(-7:7), 
     xlab = expression(bold(widehat(beta))),  cex.lab = 1.6, 
     main=bquote(Histogram~of ~bold(widehat(beta))~TSLS), 
     col=rgb(0,0,1,1/2))

plot.new()
lwd1 = 2.5
hist(bootbet4$t, breaks = 7000, freq = NULL, probability = F, xlim=range(-7:7), 
     xlab = expression(bold(widehat(beta))),  cex.lab = 1.6, main="", 
     col=rgb(0,0,1,1/3))
par(new=T)
hist(bootbet3$t, breaks = 25000, freq = NULL, probability = F, xlim=range(-7:7), 
     xlab = expression(bold(widehat(beta))), axes=F, alpha=0.3, cex.lab = 1.6, 
     main =bquote(Histogram~of ~bold(widehat(beta))~TSLS~vs.~GMM), col=rgb(1,0,0,1/3))
legend("topleft", legend=c("TSLS", "GMM"), lty= 1,  col = c(rgb(0,0,1,1/2), rgb(1,0,0,1/2)))

```


Summary statistics bootstrapped coefficients
```{r}
summary(bootbet4$t)
skewness(bootbet4$t)
kurtosis(bootbet4$t)
```


Confidence interval
```{r}
boot.ci(bootbet4)

bootbet4
```


===
GMM more simple regression
===

Specify database
```{r}
t5 <- as.data.frame(cbind(base2$logpgp95, base2$avexpr, base2$logem4))
```

TSLS
```{r}
TSLS2 <- ivreg(base2$logpgp95 ~  base2$avexpr | base2$logem4)
```

Normal GMM
```{r}
GMM1 <- gmm(base2$logpgp95 ~  base2$avexpr, ~  base2$logem4 , type="twoStep", 
            wmatrix = "optimal")
```

Bootstrap
```{r}
beta_function5 <- function(dat, i){
  resampled3 <<- dat[i,]
  bet4[i] <- return(gmm(resampled3[,1] ~ resampled3[,2], ~
                    resampled3[,3], type="twoStep", wmatrix =
         "optimal", optfct = "nlminb")$coefficients[2])
}

beta_function6 <- function(dat, i){
  resampled4 <- dat[i,]
  bet4[i] <- return(ivreg(resampled4[,1] ~ resampled4[,2] |
                    resampled4[,3])$coefficients[2])
}

## Look at standard errors much higher for optimal GMM as predicted by theory
bootbet5 <- boot(data = t5, beta_function5, R=5000)
bootbet6 <- boot(t5, beta_function6, R=5000)
```

Results
```{r}
bootbet5
bootbet6
```


Jack after Boot
```{r}
jack.after.boot(bootbet5, main ="GMM")
jack.after.boot(bootbet6, main ="2SLS")
```

Plot
```{r}
par(mfrow = c(1,2))
hist(bootbet5$t, breaks = 5000, xlim = range(-5:10), col = "blue")
hist(bootbet6$t, breaks = 3000, xlim = range(-5:10), col = "red")
```



